---
title: "COM(P)ADRE Corrections"
author: "Patrick Barks"
date: '2018-01-25'
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

&nbsp;

### Preliminaries

```{r, message = F, warning = F}
# load libraries
library(dplyr)
library(tidyr)
library(tibble)
library(purrr)

# load Comadre and Compadre databases
load('COMADRE_v.X.X.X.RData')
load('COMPADRE_v.X.X.X.RData')

# add row ids to metadata
comadre$metadata$Row <- 1:nrow(comadre$metadata)
compadre$metadata$Row <- 1:nrow(compadre$metadata)

# increase console width
options(width = 120)

# function to limit column width, for prettier Markdown document
Subs <- function(x, n) {
 ifelse(nchar(x) < n, x, paste0(substr(x, 1, n-2), "\u2026")) 
}
```

&nbsp;

### `DOI.ISBN` with >1 unique `Authors`, `YearPublication`, or `Journal`

A given DOI or ISBN should be associated with a single publication (e.g a single
journal, year, and set of authors). The code chunk below finds instances where
this is not the case. (Note: some of these cases are due to subtle differences
in spelling or formatting of author names.)

```{r}
# COMADRE
comadre$metadata %>% 
  filter(!is.na(DOI.ISBN)) %>% 
  group_by(DOI.ISBN) %>% 
  mutate(n_uniq_author = length(unique(Authors)),
         n_uniq_year = length(unique(YearPublication)),
         n_uniq_journal = length(unique(Journal))) %>% 
  ungroup() %>% 
  filter(n_uniq_author > 1 | n_uniq_year > 1 | n_uniq_journal > 1) %>% 
  select(DOI.ISBN, Journal, Year = YearPublication, Authors) %>% 
  unique() %>% 
  arrange(DOI.ISBN, Authors) %>% 
  print(n = 'all', width = 100)
```

```{r}
# COMPADRE
compadre$metadata %>% 
  filter(!is.na(DOI.ISBN)) %>% 
  group_by(DOI.ISBN) %>% 
  mutate(n_uniq_author = length(unique(Authors)),
         n_uniq_year = length(unique(YearPublication)),
         n_uniq_journal = length(unique(Journal))) %>% 
  ungroup() %>% 
  filter(n_uniq_author > 1 | n_uniq_year > 1 | n_uniq_journal > 1) %>% 
  select(DOI.ISBN, Journal, Year = YearPublication, Authors) %>% 
  unique() %>% 
  arrange(DOI.ISBN, Authors) %>%
  print(n = 'all', width = 100)
```

&nbsp;

### `Authors`x`YearPublication`x`Journal` with >1 unique `DOI.ISBN`

A given combination of author, year, and journal should generally correspond to
a single DOI or ISBN (though legitimate exceptions are possible). The code chunk
below finds instances where this is not the case.

```{r}
# COMADRE
comadre$metadata %>% 
  mutate(AuthorJournalYear = paste(Authors, Journal, YearPublication)) %>% 
  group_by(AuthorJournalYear) %>% 
  mutate(n = length(unique(DOI.ISBN))) %>% 
  ungroup() %>% 
  filter(n > 1) %>% 
  select(Authors, Year = YearPublication, Journal, DOI.ISBN) %>% 
  unique() %>% 
  arrange(Authors) %>% 
  mutate(Authors = Subs(Authors, 30))
```

```{r}
# COMPADRE
compadre$metadata %>% 
  mutate(AuthorJournalYear = paste(Authors, Journal, YearPublication)) %>% 
  group_by(AuthorJournalYear) %>% 
  mutate(n = length(unique(DOI.ISBN))) %>% 
  ungroup() %>% 
  filter(n > 1) %>% 
  select(Authors, Year = YearPublication, Journal, DOI.ISBN) %>% 
  unique() %>% 
  arrange(Authors) %>% 
  mutate(Authors = Subs(Authors, 30))
```

&nbsp;

### Get CrossRef citation info associated with each unique DOI

The code chunk below obtains citation info (journal title and first author) for
each unique DOI in COM(P)ADRE using the [CrossRef
API](https://github.com/CrossRef/rest-api-doc) (accessed via the R library
[rcrossref](https://github.com/ropensci/rcrossref)).

```{r}
# # load rcrossref
# library(rcrossref)
# 
# # function to get citation info for a given DOI, using rcrossref library
# GetCitation <- function(doi) {
#   # get citation info using rcrossref::cr_cn()
#   ref <- try(cr_cn(dois = doi, format = "citeproc-json"))
#   
#   if(is.null(ref) | class(ref) == "try-error") {
#   # if no citation found, return NAs
#     out <- data.frame(rcross_title = NA,
#                       rcross_author = NA,
#                       rcross_found = FALSE,
#                       stringsAsFactors = F)
#   } else {
#   # else, return title and author info (or NA if those elements not found)
#     rcross_title <- try(ref$`container-title`)
#     rcross_author <- try(ref$author$family[1])
#     out <- data.frame(
#       rcross_title = ifelse(is.null(rcross_title), NA, rcross_title),
#       rcross_author = ifelse(is.null(rcross_author), NA, rcross_author),
#       rcross_found = TRUE,
#       stringsAsFactors = F
#     )
#   }
#   
#   return(out)
# }
# 
# # get COMADRE citations
# citation_comadre <- tibble(DOI.ISBN = sort(unique(comadre$metadata$DOI.ISBN))) %>% 
#   slice(grep('^10', DOI.ISBN)) %>% # limit to DOIs, which start with "10."
#   group_by(DOI.ISBN) %>% 
#   do(GetCitation(.$DOI.ISBN)) %>% 
#   ungroup()
# 
# # get COMPADRE citations
# citation_compadre <- tibble(DOI.ISBN = sort(unique(compadre$metadata$DOI.ISBN))) %>% 
#   slice(grep('^10', DOI.ISBN)) %>%  # limit to DOIs, which start with "10."
#   group_by(DOI.ISBN) %>% 
#   do(GetCitation(.$DOI.ISBN)) %>% 
#   ungroup()
# 
# # write to file
# write.csv(citation_comadre, 'crossref_comadre.csv', row.names = F)
# write.csv(citation_compadre, 'crossref_compadre.csv', row.names = F)
```

```{r}
# read CrossRef citation data already written to file
citation_comadre <- read.csv('crossref_comadre.csv', stringsAsFactors = F)
citation_compadre <- read.csv('crossref_compadre.csv', stringsAsFactors = F)
```

&nbsp;

### DOI not found using the CrossRef API

The code chunk below finds DOIs listed in COM(P)ADRE that cannot be found using
the CrossRef API. These likely reflect errors in the `DOI.ISBN` column.

```{r}
# COMADRE
comadre$metadata %>% 
  select(DOI.ISBN, Journal, Year = YearPublication, Authors) %>% 
  unique() %>% 
  slice(grep('^10', DOI.ISBN)) %>% 
  left_join(citation_comadre, by = 'DOI.ISBN') %>% 
  filter(rcross_found == FALSE) %>% 
  select(-rcross_author, -rcross_title) %>% 
  mutate(Authors = Subs(Authors, 30))
```

```{r}
# COMPADRE
compadre$metadata %>% 
  select(DOI.ISBN, Journal, Year = YearPublication, Authors) %>% 
  unique() %>% 
  slice(grep('^10', DOI.ISBN)) %>% 
  left_join(citation_compadre, by = 'DOI.ISBN') %>% 
  filter(rcross_found == FALSE) %>% 
  select(-rcross_author, -rcross_title) %>% 
  mutate(Authors = Subs(Authors, 30))
```

&nbsp;

### Inconsistent journal title abbreviations

The code chunk below finds instances where there are multiple unique entries in
the `Journal` column for a given CrossRef-obtained journal title. Some of these
reflect inconsistencies in journal title abbreviations, while others reflect an
error in the `DOI.ISBN` column.

```{r}
# function to calculate the number of unique entries in the Journal column,
#  and the number of unique entries in DOI.ISBN for a given Journal entry
CheckAbbrevs <- function(dat) {
  dat %>%
    mutate(n_abbrev = length(unique(Journal))) %>% 
    group_by(Journal) %>% 
    mutate(n_doi = length(unique(DOI.ISBN))) %>%
    mutate(DOI = ifelse(n_doi > 1, "mult", unique(DOI.ISBN))) %>%
    ungroup() %>% 
    select(n_abbrev, Journal, n_doi, DOI) %>% 
    unique()
}
```

```{r}
# COMADRE
comadre$metadata %>% 
  select(DOI.ISBN, Journal) %>% 
  unique() %>% 
  left_join(citation_comadre, by = 'DOI.ISBN') %>% 
  filter(rcross_found == TRUE) %>% 
  group_by(rcross_title) %>% 
  do(CheckAbbrevs(.)) %>% 
  ungroup() %>% 
  filter(n_abbrev > 2) %>% 
  select(-n_abbrev) %>% 
  mutate(rcross_title = Subs(rcross_title, 30)) %>% 
  print(n = 'all')
```

```{r}
# COMPADRE
compadre$metadata %>% 
  select(DOI.ISBN, Journal) %>% 
  unique() %>% 
  left_join(citation_compadre, by = 'DOI.ISBN') %>% 
  filter(rcross_found == TRUE) %>% 
  group_by(rcross_title) %>% 
  do(CheckAbbrevs(.)) %>% 
  ungroup() %>% 
  filter(n_abbrev > 2) %>% 
  select(-n_abbrev) %>% 
  mutate(rcross_title = Subs(rcross_title, 30)) %>% 
  print(n = 'all')
```

&nbsp;

### Irregular formatting of author names

The code chunk below finds instances of irregular formatting in the `Authors`
column (e.g. multiple sequential spaces, or commas where there should be
semicolons).

```{r}
# check for instances where Authors column contains 2+ spaces in sequence,
#  semicolons followed by character (with no separating space), or commas
#  where there should be semicolons
irregular <- '[[:space:]][[:space:]]+|\\;[[:alpha:]]+|\\,+'
```

```{r}
# COMADRE
comadre$metadata %>% 
  select(DOI.ISBN, Journal, Year = YearPublication, Authors) %>% 
  unique() %>% 
  mutate(irreg = grepl(irregular, Authors)) %>% 
  filter(irreg == TRUE) %>% 
  select(-irreg) %>% 
  arrange(Authors) %>% 
  as_tibble() %>% 
  print(width = 100)
```

```{r}
# COMPADRE
compadre$metadata %>% 
  select(DOI.ISBN, Journal, Year = YearPublication, Authors) %>% 
  unique() %>% 
  mutate(irreg = grepl(irregular, Authors)) %>% 
  filter(irreg == TRUE) %>% 
  select(-irreg) %>% 
  arrange(Authors) %>% 
  as_tibble() %>% 
  print(width = 100)
```

&nbsp;

### Country codes different from ISO 3 standard

Country codes listed in the column `Country` should conform to the ISO 3
standard. The code chunk below finds instances where this is not the case.

```{r}
# Read list of ISO 3 country codes; obtained manually from:
# https://unstats.un.org/unsd/tradekb/knowledgebase/country-code
country_codes <- read.csv('country_codes.csv')
```

```{r}
# function to spread entries containing multiple country codes (separated by ;)
#  across separate lines
SplitCountries <- function(x) {
  data.frame(Country = strsplit(as.character(x), '; ')[[1]], stringsAsFactors = F)
}
```

```{r}
# COMADRE
comadre$metadata %>% 
  group_by(Row) %>% 
  do(SplitCountries(.$Country)) %>% 
  ungroup() %>% 
  left_join(select(comadre$metadata, Row, DOI.ISBN, Journal, Authors), by = 'Row') %>% 
  select(-Row) %>% 
  filter(!Country %in% country_codes$ISO, !is.na(Country)) %>% 
  unique() %>% 
  print(width = 100)
```

```{r}
# COMPADRE
compadre$metadata %>% 
  group_by(Row) %>% 
  do(SplitCountries(.$Country)) %>% 
  ungroup() %>% 
  left_join(select(compadre$metadata, Row, DOI.ISBN, Journal, Authors), by = 'Row') %>% 
  select(-Row) %>% 
  filter(!Country %in% country_codes$ISO, !is.na(Country)) %>% 
  unique() %>%
  print(width = 100)
```

&nbsp;

### Check for nonstandard values of `MatrixClassOrganized`

Values of `MatrixClassOrganized` should be one of 'active', 'dorm', or 'prop'
(or possibly NA). The code chunk below finds instances where this is not the
case.

```{r}
# function to check whether MatrixClassOrganized column contains 
#  values other than 'active', 'dorm', 'prop', or <NA>
MatrixClassCheck <- function (x) {
  any(!x$MatrixClassOrganized %in% c(NA, 'active', 'dorm', 'prop'))
}
```

```{r}
# COMADRE
tibble(DOI.ISBN = comadre$metadata$DOI.ISBN,
       Year = comadre$metadata$YearPublication,
       Authors = comadre$metadata$Authors,
       matrixClass = comadre$matrixClass) %>%
  mutate(flag = map_lgl(matrixClass, MatrixClassCheck)) %>% 
  filter(flag == TRUE) %>% 
  unnest() %>% 
  select(-flag, -MatrixClassAuthor) %>% 
  mutate(Authors = Subs(Authors, 30)) %>%  
  mutate(MatrixClassOrganized = Subs(MatrixClassOrganized, 20)) %>% 
  print(n = 'all')
```

```{r}
# COMPADRE
tibble(DOI.ISBN = compadre$metadata$DOI.ISBN,
       Year = compadre$metadata$YearPublication,
       Authors = compadre$metadata$Authors,
       matrixClass = compadre$matrixClass) %>%
  mutate(flag = map_lgl(matrixClass, MatrixClassCheck)) %>% 
  filter(flag == TRUE) %>% 
  unnest() %>% 
  select(-flag, -MatrixClassAuthor) %>% 
  mutate(Authors = Subs(Authors, 30)) %>% 
  mutate(MatrixClassOrganized = Subs(MatrixClassOrganized, 20)) %>% 
  print(n = 'all')
```
