---
title: "COM(P)ADRE Corrections"
author: "Patrick Barks"
date: '2018-01-25'
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

&nbsp;

### Preliminaries

```{r, message = F}
# load libraries
library(dplyr)
library(tidyr)
library(tibble)
library(purrr)

# load Comadre and Compadre databases
load('COMADRE_v.X.X.X.RData')
load('COMPADRE_v.X.X.X.RData')

# add row ids to metadata
comadre$metadata$Row <- 1:nrow(comadre$metadata)
compadre$metadata$Row <- 1:nrow(compadre$metadata)

# increase console width
options(width = 120)
```

&nbsp;

### `DOI.ISBN` with >1 unique `Authors`, `YearPublication`, or `Journal`

A given DOI or ISBN should be associated with a single publication (e.g a single
journal, year, and set of authors). The code chunk below finds instances where
this is not the case. (Note: some of these cases are due to subtle differences
in spelling or formatting of author names.)

```{r}
# COMADRE
comadre$metadata %>% 
  filter(!is.na(DOI.ISBN)) %>% 
  group_by(DOI.ISBN) %>% 
  mutate(n_uniq_author = length(unique(Authors)),
         n_uniq_year = length(unique(YearPublication)),
         n_uniq_journal = length(unique(Journal))) %>% 
  ungroup() %>% 
  filter(n_uniq_author > 1 | n_uniq_year > 1 | n_uniq_journal > 1) %>% 
  dplyr::select(DOI.ISBN, Journal, Year = YearPublication, Authors) %>% 
  unique() %>% 
  arrange(DOI.ISBN, Authors) %>% 
  mutate(Authors = substr(Authors, 1, 40)) %>% 
  as.data.frame()
```

```{r}
# COMPADRE
compadre$metadata %>% 
  filter(!is.na(DOI.ISBN)) %>% 
  group_by(DOI.ISBN) %>% 
  mutate(n_uniq_author = length(unique(Authors)),
         n_uniq_year = length(unique(YearPublication)),
         n_uniq_journal = length(unique(Journal))) %>% 
  ungroup() %>% 
  filter(n_uniq_author > 1 | n_uniq_year > 1 | n_uniq_journal > 1) %>% 
  dplyr::select(DOI.ISBN, Journal, Year = YearPublication, Authors) %>% 
  unique() %>% 
  arrange(DOI.ISBN, Authors) %>%
  mutate(Authors = substr(Authors, 1, 25)) %>% 
  as.data.frame()
```

&nbsp;

### `Authors`x`YearPublication`x`Journal` with >1 unique `DOI.ISBN`

A given combination of author, year, and journal should generally correspond to
a single DOI or ISBN (though legitimate exceptions are possible). The code chunk
below finds instances where this is not the case.

```{r}
# COMADRE
comadre$metadata %>% 
  mutate(AuthorJournalYear = paste(Authors, Journal, YearPublication)) %>% 
  group_by(AuthorJournalYear) %>% 
  mutate(n = length(unique(DOI.ISBN))) %>% 
  ungroup() %>% 
  filter(n > 1) %>% 
  dplyr::select(Authors, Year = YearPublication, Journal, DOI.ISBN) %>% 
  unique() %>% 
  arrange(Authors) %>% 
  mutate(Authors = substr(Authors, 1, 35)) %>% 
  as.data.frame()
```

```{r}
# COMPADRE
compadre$metadata %>% 
  mutate(AuthorJournalYear = paste(Authors, Journal, YearPublication)) %>% 
  group_by(AuthorJournalYear) %>% 
  mutate(n = length(unique(DOI.ISBN))) %>% 
  ungroup() %>% 
  filter(n > 1) %>% 
  dplyr::select(Authors, Year = YearPublication, Journal, DOI.ISBN) %>% 
  unique() %>% 
  arrange(Authors) %>% 
  mutate(Authors = substr(Authors, 1, 35)) %>% 
  as.data.frame()
```

&nbsp;

### Irregular formatting of author names

The code chunk below finds instances of irregular formatting in the `Authors`
column (e.g. multiple sequential spaces, or commas where there should be
semicolons).

```{r}
# check for instances where Authors column contains 2+ spaces in sequence,
#  semicolons followed by character (with no separating space), or commas
#  where there should be semicolons
irregular <- '[[:space:]][[:space:]]+|\\;[[:alpha:]]+|\\,+'
```

```{r}
# COMADRE
comadre$metadata %>% 
  dplyr::select(DOI.ISBN, Journal, Year = YearPublication, Authors) %>% 
  unique() %>% 
  mutate(irreg = grepl(irregular, Authors)) %>% 
  filter(irreg == TRUE) %>% 
  dplyr::select(-irreg) %>% 
  arrange(Authors) %>% 
  mutate(Authors = substr(Authors, 1, 40))
```

```{r}
# COMPADRE
compadre$metadata %>% 
  dplyr::select(DOI.ISBN, Journal, Year = YearPublication, Authors) %>% 
  unique() %>% 
  mutate(irreg = grepl(irregular, Authors)) %>% 
  filter(irreg == TRUE) %>% 
  dplyr::select(-irreg) %>% 
  arrange(Authors) %>% 
  mutate(Authors = substr(Authors, 1, 40))
```

&nbsp;

### Country codes different from ISO 3 standard

Country codes listed in the column `Country` should conform to the ISO 3
standard. The code chunk below finds instances where this is not the case.

```{r}
# Read list of ISO 3 country codes; obtained manually from:
# https://unstats.un.org/unsd/tradekb/knowledgebase/country-code
country_codes <- read.csv('country_codes.csv')
```

```{r}
# function to spread entries containing multiple country codes (separated by ;)
#  across separate lines
SplitCountries <- function(x) {
  data.frame(Country = strsplit(as.character(x), '; ')[[1]], stringsAsFactors = F)
}
```

```{r}
# COMADRE
comadre$metadata %>% 
  group_by(Row) %>% 
  do(SplitCountries(.$Country)) %>% 
  ungroup() %>% 
  left_join(dplyr::select(comadre$metadata, Row, DOI.ISBN, Journal, Authors), by = 'Row') %>% 
  dplyr::select(-Row) %>% 
  filter(!Country %in% country_codes$ISO, !is.na(Country)) %>% 
  unique() %>% 
  mutate(Authors = substr(Authors, 1, 25)) %>% 
  as.data.frame()
```

```{r}
# COMPADRE
compadre$metadata %>% 
  group_by(Row) %>% 
  do(SplitCountries(.$Country)) %>% 
  ungroup() %>% 
  left_join(dplyr::select(compadre$metadata, Row, DOI.ISBN, Journal, Authors), by = 'Row') %>% 
  dplyr::select(-Row) %>% 
  filter(!Country %in% country_codes$ISO, !is.na(Country)) %>% 
  unique() %>%
  mutate(Authors = substr(Authors, 1, 35)) %>%
  as.data.frame()
```

&nbsp;

### Check for nonstandard values of `MatrixClassOrganized`

Values of `MatrixClassOrganized` should be one of 'active', 'dorm', or 'prop'
(or possibly NA). The code chunk below finds instances where this is not the
case.

```{r}
# function to check whether MatrixClassOrganized column contains 
#  values other than 'active', 'dorm', 'prop', or <NA>
MatrixClassCheck <- function (x) {
  any(!x$MatrixClassOrganized %in% c(NA, 'active', 'dorm', 'prop'))
}
```

```{r}
# COMADRE
tibble(DOI.ISBN = comadre$metadata$DOI.ISBN,
       Year = comadre$metadata$YearPublication,
       Authors = comadre$metadata$Authors,
       matrixClass = comadre$matrixClass) %>%
  mutate(flag = map_lgl(matrixClass, MatrixClassCheck)) %>% 
  filter(flag == TRUE) %>% 
  unnest() %>% 
  dplyr::select(-flag, -MatrixClassAuthor) %>% 
  mutate(Authors = substr(Authors, 1, 25)) %>% 
  mutate(MatrixClassOrganized = substr(MatrixClassOrganized, 1, 20)) %>% 
  as.data.frame()
```

```{r}
# COMPADRE
tibble(DOI.ISBN = compadre$metadata$DOI.ISBN,
       Year = compadre$metadata$YearPublication,
       Authors = compadre$metadata$Authors,
       matrixClass = compadre$matrixClass) %>%
  mutate(flag = map_lgl(matrixClass, MatrixClassCheck)) %>% 
  filter(flag == TRUE) %>% 
  unnest() %>% 
  dplyr::select(-flag, -MatrixClassAuthor) %>% 
  mutate(Authors = substr(Authors, 1, 25)) %>% 
  mutate(MatrixClassOrganized = substr(MatrixClassOrganized, 1, 20)) %>% 
  as.data.frame()
```
