---
title: "Compadre Corrections"
author: "Patrick Barks"
date: '2018-01-25'
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

&nbsp;

### Load libraries and Com(p)adre databases, and adjust console width

```{r, message = F}
# load libraries
library(dplyr)
library(tidyr)
library(tibble)
library(purrr)

# load Comadre and Compadre databases
load('COMADRE_v.X.X.X.RData')
load('COMPADRE_v.X.X.X.RData')

# increase console width
options(width = 120)
```

&nbsp;

### `DOI.ISBN` with >1 unique `Authors`, `YearPublication`, or `Journal`

A given DOI or ISBN should be associated with a single publication (e.g a single
journal, year, and set of authors). The code chunk below finds instances where
this is not the case. Some of these cases seem to reflect data entry errors, but
some represent subtle differences in spelling or formatting of author names
(e.g. spelled with vs. without accents, only the first author is listed vs. all
authors listed, etc.). Also, some discrepencies in `YearPublication` for a given
DOI may reflect the time between an 'early-online' version and the final
published version.

```{r}
# COMADRE
comadre$metadata %>% 
  filter(!is.na(DOI.ISBN)) %>% 
  group_by(DOI.ISBN) %>% 
  mutate(n_uniq_author = length(unique(Authors)),
         n_uniq_year = length(unique(YearPublication)),
         n_uniq_journal = length(unique(Journal))) %>% 
  ungroup() %>% 
  filter(n_uniq_author > 1 | n_uniq_year > 1 | n_uniq_journal > 1) %>% 
  dplyr::select(DOI.ISBN, Journal, Year = YearPublication, Authors) %>% 
  unique() %>% 
  arrange(DOI.ISBN, Authors) %>% 
  mutate(Authors = substr(Authors, 1, 40)) %>% 
  as.data.frame()
```

```{r}
# COMPADRE
compadre$metadata %>% 
  filter(!is.na(DOI.ISBN)) %>% 
  group_by(DOI.ISBN) %>% 
  mutate(n_uniq_author = length(unique(Authors)),
         n_uniq_year = length(unique(YearPublication)),
         n_uniq_journal = length(unique(Journal))) %>% 
  ungroup() %>% 
  filter(n_uniq_author > 1 | n_uniq_year > 1 | n_uniq_journal > 1) %>% 
  dplyr::select(DOI.ISBN, Journal, Year = YearPublication, Authors) %>% 
  unique() %>% 
  arrange(DOI.ISBN, Authors) %>% 
  mutate(Authors = substr(Authors, 1, 25)) %>% 
  as.data.frame()
```

&nbsp;

### `Authors`x`YearPublication`x`Journal` with >1 unique `DOI.ISBN`

A given combination of author, year, and journal should generally correspond to
a single DOI or ISBN (though of course a given set of authors could publish more
than one paper in a given journal and year). Below are instances where this is
not the case. Some may reflect publications legitimately having multiple DOIs
for different versions (e.g. 'early-online version' vs. published version vs.
corrected version), though others seem to reflect data entry errors.

```{r}
# COMADRE
comadre$metadata %>% 
  mutate(AuthorJournalYear = paste(Authors, Journal, YearPublication)) %>% 
  group_by(AuthorJournalYear) %>% 
  mutate(n = length(unique(DOI.ISBN))) %>% 
  ungroup() %>% 
  filter(n > 1) %>% 
  dplyr::select(Authors, Year = YearPublication, Journal, DOI.ISBN) %>% 
  unique() %>% 
  arrange(Authors) %>% 
  mutate(Authors = substr(Authors, 1, 35)) %>% 
  as.data.frame()
```

```{r}
# COMPADRE
compadre$metadata %>% 
  mutate(AuthorJournalYear = paste(Authors, Journal, YearPublication)) %>% 
  group_by(AuthorJournalYear) %>% 
  mutate(n = length(unique(DOI.ISBN))) %>% 
  ungroup() %>% 
  filter(n > 1) %>% 
  dplyr::select(Authors, Year = YearPublication, Journal, DOI.ISBN) %>% 
  unique() %>% 
  arrange(Authors) %>% 
  mutate(Authors = substr(Authors, 1, 35)) %>% 
  as.data.frame()
```

&nbsp;

### Irregular formatting of author names

The code chunk below finds instances of irregular formatting in the `Authors`
column (e.g. multiple sequential spaces, or commas where there should be
semicolons).

```{r}
# check for instances where Authors column contains 2+ spaces in sequence,
#  semicolons followed by character (with no separating space), or commas
#  where there should be semicolons
irregular <- '[[:space:]][[:space:]]+|\\;[[:alpha:]]+|\\,+'
```

```{r}
# COMADRE
comadre$metadata %>% 
  dplyr::select(DOI.ISBN, Journal, Year = YearPublication, Authors) %>% 
  unique() %>% 
  mutate(irreg = grepl(irregular, Authors)) %>% 
  filter(irreg == TRUE) %>% 
  dplyr::select(-irreg) %>% 
  arrange(Authors) %>% 
  mutate(Authors = substr(Authors, 1, 40))
```

```{r}
# COMPADRE
compadre$metadata %>% 
  dplyr::select(DOI.ISBN, Journal, Year = YearPublication, Authors) %>% 
  unique() %>% 
  mutate(irreg = grepl(irregular, Authors)) %>% 
  filter(irreg == TRUE) %>% 
  dplyr::select(-irreg) %>% 
  arrange(Authors) %>% 
  mutate(Authors = substr(Authors, 1, 40))
```

&nbsp;

### Country codes different from ISO 3 standard

Country codes listed in the column `Country` should conform to the ISO 3
standard. The code chunk below finds instances where this is not the case.

```{r}
# Read list of ISO 3 country codes; obtained manually from:
# https://unstats.un.org/unsd/tradekb/knowledgebase/country-code
country_codes <- read.csv('country_codes.csv')
```

```{r}
# function to spread entries containing multiple country codes (separated by ;)
#  across separate lines
SplitCountries <- function(x, DOI.ISBN, Journal, Year, Authors) {
  tibble(Country = strsplit(as.character(x), '; ')[[1]],
         DOI.ISBN, Journal, Year, Authors)
}
```

```{r}
# COMADRE
comadre$metadata %>% 
  rowwise() %>% 
  do(SplitCountries(.$Country, .$DOI.ISBN, .$Journal,
                    .$YearPublication, .$Authors)) %>% 
  ungroup() %>% unique() %>% 
  filter(!Country %in% country_codes$ISO, !is.na(Country)) %>% 
  mutate(Authors = substr(Authors, 1, 20)) %>% 
  as.data.frame()
```

```{r}
# COMPADRE
compadre$metadata %>% 
  rowwise() %>% 
  do(SplitCountries(.$Country, .$DOI.ISBN, .$Journal,
                    .$YearPublication, .$Authors)) %>% 
  ungroup() %>% unique() %>% 
  filter(!Country %in% country_codes$ISO, !is.na(Country)) %>% 
  mutate(Authors = substr(Authors, 1, 35)) %>% 
  as.data.frame()
```

&nbsp;

### Check for nonstandard values of `MatrixClassOrganized`

Values of `MatrixClassOrganized` should be one of 'active', 'dorm', or 'prop'
(or possibly NA). The code chunk below finds instances where this is not the
case.

```{r}
# function to check whether MatrixClassOrganized column contains 
#  values other than 'active', 'dorm', 'prop', or <NA>
MatrixClassCheck <- function (x) {
  any(!x$MatrixClassOrganized %in% c(NA, 'active', 'dorm', 'prop'))
}
```

```{r}
# COMADRE
tibble(DOI.ISBN = comadre$metadata$DOI.ISBN,
       Year = comadre$metadata$YearPublication,
       Authors = comadre$metadata$Authors,
       matrixClass = comadre$matrixClass) %>%
  mutate(flag = map_lgl(matrixClass, MatrixClassCheck)) %>% 
  filter(flag == TRUE) %>% 
  unnest() %>% 
  dplyr::select(-flag, -MatrixClassAuthor) %>% 
  mutate(Authors = substr(Authors, 1, 25)) %>% 
  mutate(MatrixClassOrganized = substr(MatrixClassOrganized, 1, 20)) %>% 
  as.data.frame()
```

```{r}
# COMPADRE
tibble(DOI.ISBN = compadre$metadata$DOI.ISBN,
       Year = compadre$metadata$YearPublication,
       Authors = compadre$metadata$Authors,
       matrixClass = compadre$matrixClass) %>%
  mutate(flag = map_lgl(matrixClass, MatrixClassCheck)) %>% 
  filter(flag == TRUE) %>% 
  unnest() %>% 
  dplyr::select(-flag, -MatrixClassAuthor) %>% 
  mutate(Authors = substr(Authors, 1, 25)) %>% 
  mutate(MatrixClassOrganized = substr(MatrixClassOrganized, 1, 20)) %>% 
  as.data.frame()
```
