---
title: "matA != (matU + matF + matC)"
author: "Patrick Barks"
date: '2018-02-18'
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

&nbsp;

Check for entries in COM(P)ADRE where `MatrixSplit == 'Divided'`, but `matA != (matU + matF + matC)`, even after accounting for `NA`.

### Preliminaries

```{r, message = F, warning = F}
# load libraries
library(tibble)
library(dplyr)
library(Rcompadre)

# load Comadre and Compadre databases
load('COMADRE_v.X.X.X.RData')
load('COMPADRE_v.X.X.X.RData')

# add row ids to metadata
comadre$metadata$id <- 1:nrow(comadre$metadata)
compadre$metadata$id <- 1:nrow(compadre$metadata)

# function to add three matrices (replacing NA with 0)
AddMats <- function(M1, M2, M3) {
  M1[which(is.na(M1))] <- 0
  M2[which(is.na(M2))] <- 0
  M3[which(is.na(M3))] <- 0
  out <- M1 + M2 + M3
  colnames(out) <- NULL
  return(out)
}

# function to test whether all elements of two matrices equal
CompareMats <- function(A, B, prec = 0.000001) {
  return(all(abs(A - B) < prec))
}
```

### Compadre

```{r}
# subset to divided matrices
compadre_div <- subsetDB(compadre, MatrixSplit == 'Divided')

# create separate lists for mat A, U, F, C
matA <- lapply(compadre_div$mat, function(x) x$matA)
matU <- lapply(compadre_div$mat, function(x) x$matU)
matF <- lapply(compadre_div$mat, function(x) x$matF)
matC <- lapply(compadre_div$mat, function(x) x$matC)

# matU + matF + matC
sumUFC <- mapply(AddMats, matU, matF, matC, USE.NAMES = FALSE)

# compare matA to (matU + matF + matC)
mat_compare <- mapply(CompareMats, matA, sumUFC, USE.NAMES = FALSE)
mat_mismatch <- which(mat_compare == FALSE)

# which matA != (matU + matF + matC)
mat_mismatch
```

Take a closer look at some of the mismatches:

```{r}
# choose index of non-matching matrix
i <- mat_mismatch[161]

# see which transitions don't match (indicated with FALSE)
abs(matA[[i]] - sumUFC[[i]]) < 0.000001

# compare matA to (matU + matF + matC)
matA[[i]]; sumUFC[[i]]
```

In this case, it looks like fecundity transitions from the top two rows have been duplicated in matU.

```{r}
# check individual U and F (and C, if applicable)
matU[[i]]; matF[[i]]
```

Print metadata for all non-matching matrices:

```{r}
mismatch_compadre <- compadre_div$metadata %>% 
  slice(mat_mismatch) %>% 
  dplyr::select(-id)

write.csv(mismatch_compadre, 'mismatch_matA_compadre.csv', row.names = F)
```

### Comadre

```{r}
# subset to divided matrices
comadre_div <- subsetDB(comadre, MatrixSplit == 'Divided')

# create separate lists for mat A, U, F, C
matA <- lapply(comadre_div$mat, function(x) x$matA)
matU <- lapply(comadre_div$mat, function(x) x$matU)
matF <- lapply(comadre_div$mat, function(x) x$matF)
matC <- lapply(comadre_div$mat, function(x) x$matC)

# matU + matF + matC
sumUFC <- mapply(AddMats, matU, matF, matC, USE.NAMES = FALSE)

# compare matA to (matU + matF + matC)
mat_compare <- mapply(CompareMats, matA, sumUFC, USE.NAMES = FALSE)
mat_mismatch <- which(mat_compare == FALSE)

# which matA != (matU + matF + matC)
mat_mismatch
```

We can take a closer look at some of the mismatches:

```{r}
# choose index of non-matching matrix
i <- mat_mismatch[3]

# see which transitions don't match (indicated with FALSE)
abs(matA[[i]] - sumUFC[[i]]) < 0.000001

# compare matA to (matU + matF + matC)
matA[[i]]; sumUFC[[i]]
```

In this case, the U, F, and C matrices contain all zeros, even though MatrixSplit == 'Divided'.

```{r}
# check individual U, F, and C matrices
matU[[i]]; matF[[i]]; matC[[i]]
```

Print metadata for all non-matching matrices:

```{r}
mismatch_comadre <- comadre_div$metadata %>% 
  slice(mat_mismatch) %>% 
  dplyr::select(-id)

write.csv(mismatch_comadre, 'mismatch_matA_comadre.csv', row.names = F)
```

